Naturalness in U(1) extensions of the MSSM

### 2014-12-18 paper/e6ssm-tuning-paper.tex
Edited decscription of which parameters are fixed and varied to make argument clearer. This could still do with a lot more work though.

% DH:: I'm not sure if this is worth including? I've put a disclaimer in the
% text, as well.
\begin{figure}[h]
\begin{center}
\resizebox{!}{10cm}{%
\includegraphics{mssm_higgs_tuning.png}}
\caption{Scatter plot of fine tuning in the MSSM as a function of
the lightest Higgs mass, $m_{h_1}$, for the cut-off scales (from 
bottom to top) $M_X=20$ TeV, $M_X=50$ TeV, $M_X=100$ TeV and $M_X=10^{16}$ GeV.}
\label{Fig:mssmhiggstuning}
\end{center}
\end{figure}
To illustrate this, in Fig.~\ref{Fig:mssmstoptuning} we show the
variation in fine tuning when we scan over the stop masses and mixing,
with $500 \textrm{ GeV} \leq m_{Q_3},m_{u_3}\leq 10\textrm{ TeV}$ and
$-3810\textrm{ GeV}\leq A_t\leq -20\textrm{ GeV}$, for $M_X=20$ TeV,
$50$ TeV and $100$ TeV.  The remaining parameters we fix such that at
$M_{SUSY}$ they have the values $\mu=-97.5$ GeV, $B=-84.8$ GeV,
$M_1=92.1$ GeV, $M_2=95.9$ GeV, $M_3=352$ GeV, $A_b=-117.9$ GeV,
$A_\tau=-7.8$ GeV, $m_{L_i}=400$ GeV, $m_{e_i}=204$ GeV,
$m_{Q_{1,2}}=438$ GeV, $m_{u_{1,2}}436$ GeV and $m_{d_i}=438$ GeV
($i=1,2,3$). Although we should stress that making this choice will
lead to a spectrum which is in conflict with the LHC limits, doing so
ensures that fine tuning due to the other parameters is small, so that
we avoid washing out the fine tuning impact of the stops when the
tuning is small \footnote{For models in which the spectrum is heavier,
  when the stop masses are small the fine tuning reaches a lower bound
  imposed by other heavier parameters.} as can be the case when the
stop masses are less than 1 TeV.  Note that the Higgs mass is also
allowed to vary in this scan, as shown in
Fig.~\ref{Fig:mssmhiggstuning}. This reveals the tuning problem which
people have been worrying about since the discovery of the 125 GeV
Higgs boson as we see that raising the stop masses is also pushing up
the Higgs mass, meaning that heavier Higgs masses require more fine
tuning.  However for a low value of the UV scale this tuning is not so
severe unless the stops are very heavy, and a 125 geV Higgs can be
obtained without much tuning in this unrealistic case where we have
minimised other sources of tuning.  On the other hand the tuning
becomes more severe as we increase the cuttoff such that for $M_X =
10^{16}$ GeV is increased then a lightest stop masses of 1-3 TeV can
result in a fine tuning of $\approx 200 - 1000$ as shown in
Fig.~\ref{Fig:mssmstoptuning} and the minimum tuning we see consistent with a $125$ GeV Higgs is $\approx 200$.\\ \\

@@

Since the stop mass does not have a large impact on the fine tuning when the 
cutoff scale is very low we can use this to see more clearly the impact of the
$Z^\prime$ mass on fine tuning. To do so we select a fixed low cutoff of $M_X=20$ TeV and compare the
fine tuning for two different values of the $Z^\prime$ mass. We choose to look at
$M_{Z^\prime}=2.5$ TeV, which is just above the current limits, and $M_{Z^\prime}=4.5$ TeV, 
which should be in reach in run II at the LHC. We also compare the fine tuning calculated
in each case to the tuning in the MSSM. For this, we have performed a parameter space scan in both the 
MSSM and E$_6$SSM, scanning over those parameters that appear at tree level in the EWSB conditions in the
models. While the contribution from large stop masses to the fine tuning is small for such a low cut-off scale, the 
stop contributions to the effective potential can play a significant role in reducing the fine tuning.
For this reason it is still important to properly treat the tuning associated with stop contributions to the one-loop effective potential, and so we also scan over the soft masses $m_{Q_3}^2$, $m_{u_3}^2$ and the stop mixing $A_t$.
The relevant parameters and ranges that were scanned over are summarised in Table \ref{tab:scanranges}. With the exception of the stop mixing $A_t$ and the soft trilinear $A_\lambda$, which are logarithmically scanned, all parameters are linearly scanned. In both the MSSM and E$_6$SSM, we then decouple all other parameters by setting the remaining soft masses to values of $5$ TeV, while the first and second generation Yukawa couplings and their associated soft trilinears are taken to vanish. The $U(1)$ gaugino soft mass $M_1$ was fixed to $M_1=300$ GeV, and we fix $M_3=2000$ GeV. Additionally, in the E$_6$SSM the $U(1)_N$ gaugino soft mass $M_1'$ is held fixed with $M_1'=M_1=300$ GeV, and $\mu^\prime=5$ TeV.
\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cc}
MSSM & E$_6$SSM \\
\hline
$2 \leq \tan\beta \leq 50$& $2 \leq \tan\beta \leq 50$ \\
$-1\textrm{ TeV } \leq \mu \leq 1 \textrm{ TeV}$ & $-3 \leq \lambda \leq 3$\\
$-1\textrm{ TeV } \leq B \leq 1\textrm{ TeV}$ & $-10\textrm{ TeV } \leq A_\lambda \leq 10\textrm{ TeV}$ \\
$ 200 \textrm{ GeV } \leq m_{Q_3} \leq 2000 \textrm { GeV}$ & $ 200 \textrm{ GeV } \leq m_{Q_3} \leq 2000 \textrm { GeV}$\\
$ 200 \textrm{ GeV } \leq m_{u_3} \leq 2000 \textrm { GeV}$ & $ 200 \textrm{ GeV } \leq m_{u_3} \leq 2000 \textrm { GeV}$\\
$ -10 \textrm{ TeV } \leq A_t \leq 10 \textrm { TeV}$ & $ -10 \textrm{ TeV } \leq A_t \leq 10 \textrm { TeV}$\\
$M_2=100\textrm{ GeV, } 1050\textrm{ GeV, } 2000 \textrm{ GeV}$ & $M_2=100\textrm{ GeV, } 1050\textrm{ GeV, } 2000 \textrm{ GeV}$ \\
\end{tabular}
\end{ruledtabular}
\caption{The parameters scanned over and the ranges of values used in the MSSM and the $E_6$ models.}
\label{tab:scanranges}
\end{table}

>>>>

% DH:: I'm not sure if this is worth including? I've put a disclaimer in the
% text, as well.
\begin{figure}[h]
\begin{center}
\resizebox{!}{10cm}{%
\includegraphics{mssm_higgs_tuning.png}}
\caption{Scatter plot of fine tuning in the MSSM as a function of
the lightest Higgs mass, $m_{h_1}$, for the cut-off scales (from 
bottom to top) $M_X=20$ TeV, $M_X=50$ TeV, $M_X=100$ TeV and $M_X=10^{16}$ GeV.}
\label{Fig:mssmhiggstuning}
\end{center}
\end{figure}
To illustrate this, in Fig.~\ref{Fig:mssmstoptuning} we show the
variation in fine tuning when we scan over the stop masses and mixing,
with $500 \textrm{ GeV} \leq m_{Q_3},m_{u_3}\leq 10\textrm{ TeV}$ and
$-3810\textrm{ GeV}\leq A_t\leq -20\textrm{ GeV}$, for $M_X=20$ TeV,
$50$ TeV and $100$ TeV.  The remaining parameters we fix such that at
$M_{SUSY}$ they have the values $\mu=-97.5$ GeV, $B=-84.8$ GeV,
$M_1=92.1$ GeV, $M_2=95.9$ GeV, $M_3=352$ GeV, $A_b=-117.9$ GeV,
$A_\tau=-7.8$ GeV, $m_{L_i}=400$ GeV, $m_{e_i}=204$ GeV,
$m_{Q_{1,2}}=438$ GeV, $m_{u_{1,2}}436$ GeV and $m_{d_i}=438$ GeV
($i=1,2,3$). Although we should stress that making this choice will
lead to a spectrum which is in conflict with the LHC limits, doing so
ensures that fine tuning due to the other parameters is small, so that
we avoid washing out the fine tuning impact of the stops when the
tuning is small\footnote{For models in which the spectrum is heavier,
  when the stop masses are small the fine tuning reaches a lower bound
  imposed by other heavier parameters.} as can be the case when the
stop masses are less than 1 TeV.  Note that the Higgs mass is also
allowed to vary in this scan, as shown in
Fig.~\ref{Fig:mssmhiggstuning}. This reveals the tuning problem which
people have been worrying about since the discovery of the 125 GeV
Higgs boson as we see that raising the stop masses is also pushing up
the Higgs mass, meaning that heavier Higgs masses require more fine
tuning.  However for a low value of the UV scale this tuning is not so
severe unless the stops are very heavy, and a 125 geV Higgs can be
obtained without much tuning in this unrealistic case where we have
minimised other sources of tuning.  On the other hand the tuning
becomes more severe as we increase the cuttoff such that for $M_X =
10^{16}$ GeV a lightest stop masses of $1-3$ TeV can result in a fine
tuning of $\approx 200 - 1000$ as shown in
Fig.~\ref{Fig:mssmstoptuning} and the minimum tuning we see consistent
with a $125$ GeV Higgs is $\approx 200$, as shown in
Fig.~\ref{Fig:mssmhiggstuning}.

@@

Since the stop mass does not have such a large impact on the fine
tuning when the cutoff scale is very low we can use this to see more
clearly the impact of the $Z^\prime$ mass on fine tuning. To do so we
select a fixed low cutoff of $M_X=20$ TeV and compare the fine tuning
for two different values of the $Z^\prime$ mass. We choose to look at
$M_{Z^\prime}=2.5$ TeV, which is just above the current limits, and
$M_{Z^\prime}=4.5$ TeV, which should be in reach in run II at the LHC
and then compare the fine tuning calculated in each case to the tuning
in the MSSM.  For this, we have performed a large six dimensional
parameter space scan in both the MSSM and E$_6$SSM, varying those
parameters most relevant for the fine tuning and the Higgs mass.
Therefore the set of parameters which we vary includes $\mu$, $B$ and
$\tan \beta$ for the MSSM, and $\lambda$, $A_\lambda$ and $\tan
\beta$, for the E$_6$SSM, which appear at tree level in the EWSB
conditions of the models. While the contribution from large stop
masses to the fine tuning is small for such a low cut-off scale, the
stop contributions to the effective potential can play a significant
role in reducing the fine tuning.  For this reason it is still
important to properly treat the tuning associated with stop
contributions to the one-loop effective potential, and so we also scan
over the soft masses $m_{Q_3}^2$, $m_{u_3}^2$ and the stop mixing
$A_t$.  The relevant parameters and ranges that were scanned over are
summarised in Table \ref{tab:scanranges}. With the exception of the
stop mixing $A_t$ and the soft trilinear $A_\lambda$, which are
logarithmically scanned, all parameters are linearly scanned.  In
addition to this we also repeat each scan for three different values
of $M_2$ to introduce more variation in the chargino masses, which can
also be importnat due to the link between $\mu$ and the fine tuning.

 In this case we now consider realistic scenarios, where the
 parameters that are not scanned over are set to values which keep the
 associated states comfortably above their experimental limits.  So in
 both the MSSM and E$_6$SSM, all other soft [scalar?] masses are set
 to $5$ TeV.  Since we work in the third family approximation, taking
 the first and second generation Yukawa couplings to be zero, we also
 assume their associated soft trilinears vanish [what about
   $A_\tau$?]. The $U(1)$ gaugino soft mass $M_1$ was fixed to
 $M_1=300$ GeV, and we fix $M_3=2000$ GeV. Additionally, in the
 E$_6$SSM the $U(1)_N$ gaugino soft mass $M_1'$ is held fixed with
 $M_1'=M_1=300$ GeV, and $\mu^\prime=5$ TeV.

\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cc}
MSSM & E$_6$SSM \\
\hline
$2 \leq \tan\beta \leq 50$& $2 \leq \tan\beta \leq 50$ \\
$-1\textrm{ TeV } \leq \mu \leq 1 \textrm{ TeV}$ & $-3 \leq \lambda \leq 3$\\
$-1\textrm{ TeV } \leq B \leq 1\textrm{ TeV}$ & $-10\textrm{ TeV } \leq A_\lambda \leq 10\textrm{ TeV}$ \\
$ 200 \textrm{ GeV } \leq m_{Q_3} \leq 2000 \textrm { GeV}$ & $ 200 \textrm{ GeV } \leq m_{Q_3} \leq 2000 \textrm { GeV}$\\
$ 200 \textrm{ GeV } \leq m_{u_3} \leq 2000 \textrm { GeV}$ & $ 200 \textrm{ GeV } \leq m_{u_3} \leq 2000 \textrm { GeV}$\\
$ -10 \textrm{ TeV } \leq A_t \leq 10 \textrm { TeV}$ & $ -10 \textrm{ TeV } \leq A_t \leq 10 \textrm { TeV}$\\
$M_2=100\textrm{ GeV, } 1050\textrm{ GeV, } 2000 \textrm{ GeV}$ & $M_2=100\textrm{ GeV, } 1050\textrm{ GeV, } 2000 \textrm{ GeV}$ \\
\end{tabular}
\end{ruledtabular}
\caption{The parameters scanned over and the ranges of values used in the MSSM and the $E_6SSM$ models.}
\label{tab:scanranges}
\end{table}
